Introducing statistics with `intRo`

### 2016-01-20 jcgs-paper-2016/paper.Rmd
Some work on the interpolation section

---
title: "Introducing statistics with `intRo`"
author:
  - name: Eric Hare
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.tex
    number_sections: yes
    template: jcgs_article.tex
abstract: |
  `intRo` is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`. `intRo` is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
keywords: blah, blah blah
bibliography: refs.bib
blind: false
---

@@

As an example of possible customization, an instructor may wish to introduce hypothesis testing from a nonparametric framework rather than distributional. In this case, a wilcoxon rank sum test could be added easily as a module to be used instead of the T test. This addition would be fairly simple to complete. The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `observe.R`, `output.R`, `reactive.R`, `static.R`, and `ui.R`. Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application. The `ui.R` file contains all the user interface elements within the new module's tab. This will include any options in the options panel as well as results. The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module. The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{Shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{Shiny} framework. By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application.

@@

We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an \texttt{intRo} module significantly more seamless. The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham[ref]. The idea of interpolation is that the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically. This means that depending on the options selected in \texttt{intRo}, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.

```{r, eval=FALSE}
interpolate <- function(code, ..., mydir, `_env` = parent.frame(), 
                        file = "code_All.R", append = FALSE, 
                        save_result = FALSE, eval = TRUE) {
    stopifnot(inherits(code, "formula"), length(code) == 2)

    args <- lapply(list(...), as_call)
    expr <- methods::substituteDirect(as_call(code), args)

    cat(paste0(as.character(expr)[2], "\n"), 
        file = file.path(mydir, file), append = append)

    if (save_result) cat(paste0(paste(readLines(
        file.path(mydir, file)), collapse = "\n"), "\n"), 
        file = file.path(mydir, "code_All.R"), append = TRUE)
    if (eval) eval(expr, `_env`)
}
```

@@

\texttt{intRo} will be trialed in an introductory statistics course at Iowa State University in Spring 2015. This will allow us to assess its usefulness relative to software used in the past, as well as guage areas for improvement. Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.

>>>>

---
title: "Introducing statistics with `intRo`"
author:
  - name: Eric Hare
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: preamble.tex
    number_sections: yes
    template: jcgs_article.tex
abstract: |
  `intRo` is a modern web-based application for performing basic data analysis and statistical routines. Leveraging the power of `R` and `Shiny`, `intRo` implements common statistical functions in a powerful and extensible modular structure, while remaining simple enough for the novice statistician. This simplicity lends itself to a natural presentation in an introductory statistics course as a substitute for other commonly used statistical software packages, such as `Excel` and `JMP`. `intRo` is currently deployed at the URL \url{http://www.intro-stats.com}. Within this paper, we introduce the application and explore the design decisions underlying `intRo`, as well as highlight some challenges and advantages of reactive programming.
keywords: R, Shiny, Statistics, Teaching, Interactive
bibliography: refs.bib
blind: false
---

@@

As an example of possible customization, an instructor may wish to introduce hypothesis testing from a nonparametric framework rather than distributional. In this case, a wilcoxon rank sum test could be added easily as a module to be used instead of the T test. This addition would be fairly simple to complete. The user would fork \texttt{intRo} from http://github.com/gammarama/intRo and begin by adding six types of files to a folder within a module folder: `helper.R`, `libraries.R`, `observe.R`, `output.R`, `reactive.R`, and `ui.R`. Each of these files is written to comply with the reactive nature available in a \texttt{Shiny} application. The `ui.R` file contains all the user interface elements within the new module's tab. This will include any options in the options panel as well as results. The `helper.R` and `static.R` files refer to static functions and global variable that may be used within the module. The file `output.R` contains the reactive elements that tie output to the user interface within \texttt{Shiny} and `observe.R` and `reactive.R` contain elements within an `observer` function or `reactive` function within the \texttt{Shiny} framework. By placing each of these files within the `modules/nonparametric_test` folder and adding a single line to a navigation file indicating under which module grouping the non-parametric test module should go, a new module is available within \texttt{intRo} upon running the application. Further details on the creation of new modules can be found in section \ref{modules}.

@@

We designed the modules in such a way that the code printed to the console is the same as the code executed by the server, which makes the process of writing an \texttt{intRo} module significantly more seamless. The workhorse function for this reproducibility framework is called `interpolate`, which is based on a function proposed by Hadley Wickham [-@interpolate]. The idea of interpolation is that by using non-standard evaluation, the call can be captured, printed, and evaluated, with all the appropriate arguments substituted into the call automatically. This means that depending on the options selected in \texttt{intRo}, the corresponding arguments to functions are appropriately set for reproduction in a standard R session.

@@

We hope to use \texttt{intRo} in courses to collect feedback regarding the ease of use and functionality. This will allow us to assess its usefulness relative to software used in the past, as well as gauge areas for improvement. Furthermore, we can determine the effectiveness of code printing on generating excitement from the students about programming in \texttt{R}.
